---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(stringr)
library(janitor)

rwa_data <- read_csv("raw_data/rwa.csv") %>% 
  clean_names()



rwa_data_reverse <- rwa_data %>% 
  mutate(
    across(c(q4, q6, q8, q9, q11, q13, q15, q18, q20, q21),
           ~ case_when(
             . == 0 ~ 9,
             . == 1 ~ 8,
             . == 2 ~ 7,
             . == 3 ~ 6,
             . == 4 ~ 5,
             . == 5 ~ 4,
             . == 6 ~ 3,
             . == 7 ~ 2,
             . == 8 ~ 1,
             . == 9 ~ 0)
           )
  )

rwa_data_reverse <-  rwa_data_reverse %>% 
  mutate(gender = 
           case_when(
             gender == 1 ~ "Male",
             gender == 2 ~ "Female",
             gender == 3 ~ "Other"
           )
  )

  
rwa_data_reverse %>% 
  pivot_longer(cols = c(q3:q22), 
               names_to = "question", 
               values_to = "score") %>% 
  select(question, score, gender) %>% 
  group_by(gender) %>% 
  drop_na() %>% 
  summarise(avg_score = mean(score))
  

```

## What’s the average RWA score for left handed people vs. right handed people 

```{r}

rwa_data_reverse %>% 
  mutate(hand = 
           case_when(
    hand == 1 ~ "Right",
    hand == 2 ~ "Left",
    hand == 3 ~ "Both"
  )
    
  ) %>% 
   pivot_longer(cols = c(q3:q22), 
               names_to = "question", 
               values_to = "score") %>% 
  select(question, score, hand) %>% 
  group_by(hand) %>% 
  drop_na() %>% 
  summarise(avg_score = mean(score))
  
```

## What’s the average family size for each type of childhood?

```{r}
rwa_data_reverse %>% 
  mutate(urban = 
           case_when(
             urban == 1 ~ "Rural",
             urban == 2 ~ "Suburban",
             urban == 3 ~ "Urban"
           )) %>% 
  select(familysize, urban) %>% 
  group_by(urban) %>% 
  drop_na() %>% 
  summarise(avg_fam = mean(familysize))
```

## What’s the average time to take the test for each education level?

## 1=Less than high school, 2=High school, 3=University degree, 4=Graduate degree

```{r}
rwa_data_reverse %>% 
  mutate(education = 
           case_when(
             education == 1 ~ "Less Than High School",
             education == 2 ~ "High School",
             education == 3 ~ "University Degree",
             education == 4 ~ "Graduate Degree"
           )) %>% 
  pivot_longer(cols = c(e3:e22),
               names_to = "question_time",
               values_to = "time_taken") %>% 
  select(education, question_time, time_taken) %>% 
  group_by(education) %>% 
  drop_na() %>% 
  summarise(avg_time = mean(time_taken)) 



rm(plot_question_4)
  
```
Create a plot of results of question 4.

```{r}
rwa_data_reverse %>% 
  mutate(education = 
           case_when(
             education == 1 ~ "Less Than High School",
             education == 2 ~ "High School",
             education == 3 ~ "University Degree",
             education == 4 ~ "Graduate Degree"
           )) %>% 
  pivot_longer(cols = c(e3:e22),
               names_to = "question_time",
               values_to = "time_taken") %>% 
  select(education, question_time, time_taken) %>% 
  group_by(education) %>% 
  drop_na() %>% 
  mutate(avg_time = mean(time_taken)) %>% 
  ggplot(aes(x = education, y = avg_time)) + geom_col()

rm(plot_question_4)

```

## What’s the average RWA score for people aged
## Under 18, 18 to 25, 26 to 40, 41 to 60, Over 60

```{r}

rwa_data_reverse %>% 
  pivot_longer(cols = c(q3:q22), 
               names_to = "question", 
               values_to = "score") %>% 
  mutate(age_range = 
           case_when(
             age > 60 ~ "Over 60",
             age > 40 ~ "41 to 60",
             age > 25 ~ "26 to 40",
             age > 17 ~ "18 to 25",
             TRUE ~ "Under 18"
           )) %>% 
  select(question, score, age_range) %>% 
  group_by(age_range) %>% 
  summarise(avg_score = mean(score))

```

