---
title: "R Notebook"
output: html_notebook
---

1. The client only counts a valid email address as one ending in ‘.com’. 
How many survey results have a valid email address.

```{r}
pattern <- "\\.com"

distinct_dog %>% 
  mutate(email_detect = str_detect(email, pattern)) %>% 
  filter(email_detect == TRUE) %>% 
  summarise(total = sum(email_detect))
```

2. What’s the average amount spent on dog food for each dog size.

```{r}
pattern_spent <- "[[:alpha:]]+"
pattern_punct <- "[^.[:^punct:]]+"
pattern_double_num <- " £20$"
pattern_pound_symbol <- "£"

spent_clean <- distinct_dog %>% 
  mutate(amount_spent_on_dog_food = str_replace_all(amount_spent_on_dog_food, pattern_spent, ""),
         amount_spent_on_dog_food = str_replace_all(amount_spent_on_dog_food, pattern_punct, ""),
         amount_spent_on_dog_food = str_replace_all(amount_spent_on_dog_food, pattern_double_num, ""),
         amount_spent_on_dog_food = str_replace_all(amount_spent_on_dog_food, pattern_pound_symbol, ""),
         amount_spent_on_dog_food = as.numeric(amount_spent_on_dog_food))

spent_clean

spent_clean %>% 
  distinct(dog_size)

pattern_dog_size <- "[[:punct:]]"

clean_size_spent <- spent_clean %>% 
  mutate(dog_size = str_replace_all(dog_size, pattern_dog_size, ""),
         dog_size = replace(dog_size, dog_size == c("NO", "NA", "SLL"), NA),
         dog_size  = 
           case_when(
             dog_size == "Smallish" ~ "S",
             dog_size == "Medium sized" ~ "M",
             dog_size == "large" ~ "L",
             TRUE ~ dog_size
           ),
         dog_size = na_if(dog_size, "NA"),
         dog_size = na_if(dog_size, ""))

clean_size_spent %>% 
  select(dog_size, amount_spent_on_dog_food) %>% 
  drop_na() %>% 
  group_by(dog_size) %>% 
  summarise(avg_spend = mean(amount_spent_on_dog_food))



```

3. For owners whose surname starts with a letter in the second half of the 
alphabet (N onwards) what is the average age of their dog?

```{r}
full_clean_dog <- clean_size_spent %>% 
  mutate(dog_age = if_else(str_detect(dog_age, pattern_spent), "NA", dog_age),
         dog_age = if_else(str_detect(dog_age, pattern_punct), "NA", dog_age),
         dog_age = str_replace_all(dog_age, "[0-9]\\+$", ""),
         dog_age = na_if(dog_age, "NA"),
         dog_age = as.numeric(dog_age)) %>% 
  mutate(real_age = round(dog_age/6, digits = 0))

full_clean_dog %>% 
  select(last_name, real_age) %>% 
  filter(str_detect(last_name, "[N-Z]")) %>% 
  drop_na() %>% 
  summarise(avg_age = mean(real_age))
```

4. The dog_age column is the age in dog years. If the conversion is 1 human 
year = 6 dog years, then what is the average human age for dogs of each gender?

```{r}
full_clean_dog %>% 
  distinct(dog_gender)

pattern_female <- "[fF][eE][mM][aA]?[lL][aA]?[eE]"
pattern_male <- "^[mM][aA][lL][eE]"

cleaniest_dog <- full_clean_dog %>% 
  mutate(dog_gender = if_else(str_detect(dog_gender, pattern_dog_size), "NA", dog_gender),
         dog_gender = str_replace_all(dog_gender, pattern_female, "F"),
         dog_gender = str_replace_all(dog_gender, pattern_male, "M"),
         dog_gender = case_when(
           dog_gender == "F" ~ "Female",
           dog_gender == "M" ~ "Male",
           TRUE ~ "NA"),
         dog_gender = na_if(dog_gender, "NA")
         )

cleaniest_dog %>% 
  select(dog_gender, real_age) %>% 
  group_by(dog_gender) %>% 
  drop_na() %>% 
  summarise(avg_age = mean(real_age)) 


```

5. plot for question 4

```{r}

cleaniest_dog %>% 
  select(dog_gender, real_age) %>% 
  group_by(dog_gender) %>% 
  drop_na() %>% 
  summarise(avg_age = mean(real_age)) %>% 
  ggplot(aes(dog_gender, avg_age)) + geom_point()
```

