---
title: "R Notebook"
output: html_notebook
---

# Introduction to dataset

The Decathlon dataset looks at data for athletes and records their performance
in multiple track and field events. 

# Assumptions

1. The column names needed to be tided
2. The row names would have to be altered to a column
3. There might be missing values
4. Competition column values need a name change
5. 'Javeline' column is spelt wrong - need to change

# Cleaning the data

1. Imported libraries that I will use to clean my data with

```{r}
library(janitor)
library(tidyverse)
library(stringr)
library(readr)
```

2. Data has been read in using read_rds, checked the names of the columns and 
called the view() function to have a better look at the data

```{r}
decathlon <- read_rds("raw_data/decathlon.rds")

names(decathlon)
view(decathlon)
```

3. I then wrote a cleaning script which began with the clean_names() function 
to tidy all the names to a more readable lower and snake case. 

4. I renamed the miss spelled "Javeline" column, then created a new column 
called "names" which included all athletes' names previously held in the row 
names section.

5. I then uppercased the first letter of the values in the new "names" column

6. Finally, I created a pattern to fix "OlympicG" to "Olympics"

```{r}
pattern <- "[G]"

decathlon <- decathlon %>% 
  # clean column names
  clean_names() %>% 
  # rename "Javeline" to correct spelling of "Javelin" 
  rename(javelin = "javeline") %>% 
  # change the row names to a new column called "name"
  rownames_to_column("name") %>% 
  # mutate new column "name" to uppercase first letter of names in column
  mutate(name = str_to_title(name)) %>% 
  # use the pattern above to change value "OlympicG" to "Olympics"
  mutate(competition = str_replace_all(competition, pattern, "s"))
```


7. I saved the clean data to clean_data folder as a csv file
```{r}
write_csv(decathlon, "clean_data/decathlon_clean.csv")
```


# Data Analysis

```{r}
library(tidyverse)

library(here)

here::here()

decathlon <- read_csv(here("clean_data/decathlon_clean.csv"))
names(decathlon)

decathlon
```

# Who has the longest long jump seen in the data?

```{r}

decathlon %>% 
  #select columns I need
  select(name, long_jump) %>% 
  #find the max value for long jump and associated name
  slice_max(long_jump)

# Clay - 7.96m
```

# What was the average 100m time in each competition?

```{r}

decathlon %>% 
  # select the columns I need
  select(x100m, competition) %>% 
  # group by competition 
  group_by(competition) %>% 
  # find the average for each competition by calculating the mean of 100m times
  summarise(avg_100m = mean(x100m))

# Decastar - 11.17538
# Olympics - 10.91571
```

# Who had the highest total points across both competitions? 

```{r}

decathlon %>% 
  # select columns I need
  select(name, points) %>% 
  # group by names
  group_by(name) %>% 
  # get the total points for each person across both competitions
  summarise(total_points = sum(points)) %>% 
  # return the top athlete by total points 
  slice_max(total_points)


# Sebrle - 17110
```

# What was the shot-put scores for the top three competitors in each competition?

```{r}

decathlon %>% 
  # return rows for only the Decastar competition
  filter(competition == "Decastar") %>% 
  # select name and shot put scores
  select(name, shot_put) %>% 
  # arrange so top score is first
  arrange(desc(shot_put)) %>% 
  # return the first three values
  head(3)

# Yurkov	15.19			
# Sebrle	14.83			
# # Karpov	14.77	

# repeated the same as above except now looking at the Olympics
decathlon %>% 
  filter(competition == "Olympics") %>% 
  select(name, shot_put) %>% 
  arrange(desc(shot_put)) %>% 
  head(3)

# Sebrle	16.36			
# Karpov	15.93			
# Macey	15.73	
```

# What was the average points for competitors who ran the 400m in less than 
# 50 seconds vs. those who ran the 400m in more than 50 seconds? 

```{r}
decathlon %>% 
  # created a new logic column that states whether they ran over or under 50 secs
  mutate(under_or_over_50 = if_else(x400m < 50, "ran under 50 secs", "ran over 50 secs")) %>%
  # grouped by the new column
  group_by(under_or_over_50) %>% 
  # summarised the avg_points for those who ran under and those who ran over 50 secs
  summarise(avg_points = mean(points))

# ran over 50 secs	7727.167			
# ran under 50 secs	8120.483	
```

