---
title: "R Notebook"
output: html_notebook
---

```{r}
clean_candy <- read_csv(here::here("clean_data/clean_candy_data.csv"))

clean_candy
```

## What is the total number of candy ratings given across the three years. 
## (Number of candy ratings, not the number of raters. Don’t count missing values)

```{r}
view(clean_candy)

clean_candy %>% 
   summarise(total_reviews = 
    sum(across(
      .cols = butterfinger:take_5,
      .fns = ~sum(!is.na(.x))
    ))
  )

 
```

## What was the average age of people who are going out trick or treating?

```{r}
clean_candy %>% 
  select(age, going_out) %>% 
  filter(going_out == "Yes") %>% 
  drop_na(age) %>% 
  summarise(avg_age = mean(age))
```

## What was the average age of people who are not going trick or treating?

```{r}
clean_candy %>% 
  select(age, going_out) %>% 
  filter(going_out == "No") %>% 
  drop_na(age) %>% 
  summarise(avg_age = mean(age))
```

## For each of joy, despair and meh, which candy bar received the most of these 
## ratings?

```{r}

clean_candy %>% 
   pivot_longer(cols = butterfinger:take_5,
               names_to = "candy",
               values_to = "rating") %>% 
  select(candy, rating) %>% 
  group_by(candy, rating) %>% 
  drop_na() %>% 
  summarise(total = n()) %>% 
  ungroup() %>% 
  group_by(rating) %>% 
  slice_max(total)
  
```

## How many people rated Starburst as despair?

```{r}
clean_candy %>% 
   pivot_longer(cols = butterfinger:take_5,
               names_to = "candy",
               values_to = "rating") %>% 
  select(candy, rating) %>% 
  filter(candy == "starburst") %>% 
  group_by(rating) %>% 
  summarise(n = n()) %>% 
  ungroup() %>% 
  filter(rating == "DESPAIR")
  
```

## What was the most popular candy bar by this rating system for each gender in 
## the dataset ?

## despair -1, joy +1, meh 0 

```{r}

clean_candy %>% 
   pivot_longer(cols = butterfinger:take_5,
               names_to = "candy",
               values_to = "rating") %>% 
  mutate(rating_system = case_when(
    rating == "DESPAIR" ~ -1,
    rating == "JOY" ~ 1,
    rating == "MEH" ~ 0
  )) %>% 
  select(candy, gender, rating_system) %>% 
  drop_na() %>% 
  group_by(gender, candy) %>% 
  summarise(total = sum(rating_system)) %>% 
  ungroup() %>% 
  group_by(gender) %>% 
  slice_max(total)
  

```
## What was the most popular candy bar in each year?

```{r}

clean_candy %>% 
   pivot_longer(cols = butterfinger:take_5,
               names_to = "candy",
               values_to = "rating") %>% 
  mutate(rating_system = case_when(
    rating == "DESPAIR" ~ -1,
    rating == "JOY" ~ 1,
    rating == "MEH" ~ 0
  )) %>% 
  select(year, candy, rating_system) %>% 
  drop_na() %>% 
  group_by(candy, year) %>% 
  summarise(total = sum(rating_system)) %>% 
  ungroup() %>% 
  group_by(year) %>% 
  slice_max(total)
  
  

```

## What was the most popular candy bar by this rating for people in US, Canada, 
## UK, and all other countries?

```{r}

clean_candy %>% 
   pivot_longer(cols = butterfinger:take_5,
               names_to = "candy",
               values_to = "rating") %>% 
  mutate(rating_system = case_when(
    rating == "DESPAIR" ~ -1,
    rating == "JOY" ~ 1,
    rating == "MEH" ~ 0
  )) %>% 
  select(country, candy, rating_system) %>% 
  drop_na() %>% 
  group_by(country, candy) %>% 
  summarise(total = sum(rating_system)) %>% 
  ungroup() %>% 
  group_by(country) %>% 
  slice_max(total)

```

