---
title: "Untitled"
author: "Conal Henderson"
date: "2022-10-26"
output: html_document
---

## A brief introduction to the dataset

The data analysed for this project were the "boing-boing-candy" datasets which consisted of three seperate tables each representing consumer ratings for over 100 differnet candies from 2015, 2016 and 2017. Each dataset was extremely messy and unstructured in the sense that they could not easily be joined without first cleaning each individually. Once cleaned to a respectable standard, the datasets were joined to begin the final cleaning stage which would provide the basis of my analysis. 

## A list of any assumptions you have made

1. All non-candy data could be dropped
2. Since there was no primary and foreign keys, I could not perform a join
3. Column names would have to be renamed
4. I needed to unite columns together from each dataset
5. A column to represent the year would have to be created
6. Key data which identified age, country, state would have to be cleaned to conduct proper analysis

## The steps you took to clean the data (you don’t need to write out in detail every step, but a combination of commented code chunks with surrounding overall explanations would be great).

1. I loaded in the data, checked each dataset individudally to see what I could do straight away. 
I then proceded to write a function that would clean column names and remove any empty rows and columns. 

```{r}
clean_tables <- function(data_input){ # takes in a data set
  data_input %>% 
    # cleans names
    clean_names() %>% 
    # removes empty rows and columns
    remove_empty(which = c("rows", "cols"))
}
```

2. I then used the compare_df_cols() function from the janitor library which allowed me to compare which columns matched and were, therefore, ready to be joined. I first compared the data from 2015 and 2016 with the intuition that once they had been joined I would then match their columns with the columns in 2017. By setting the bind_method to "rbind" and return to "mismatch" it allowed me to see the that could not be bound and would require column name changes to carry out a bind. 

```{r}
compare_df_cols(candy_2015, candy_2016, return = "mismatch", bind_method = "rbind")
```

3. Once identified some columns that would easily be changed before binding rows, I wrote a short function to rename them. 

```{r}
change_names <- function(data_input){ # takes data input
  data_input %>% 
    # rename the following columns present in each data set
    rename(going_out = "are_you_going_actually_going_trick_or_treating_yourself", 
           age = "how_old_are_you",
           brown_globs = "anonymous_brown_globs_that_come_in_black_and_orange_wrappers",
           )
}
```


4. I identified some columns that could not be altered by a general function, once they were renamed it allowed for a smoother process when I came to bind the datasets together. 


```{r}
candy_2016 <- candy_2016 %>% 
  rename(bonkers = "bonkers_the_candy"
         )

candy_2015 <- candy_2015 %>% 
  rename(boxo_raisins = "box_o_raisins",
         hersheys_dark_chocolate = "dark_chocolate_hershey",
         hersheys_kisses = "hershey_s_kissables",
         )
```


5. Once I felt each dataset aligned well, I then bound them together and saved 
them as a new variable 

```{r}
data_2015_2016 <- bind_rows(candy_2015, candy_2016)
```

6. I carried out the same process as before, starting by using the compare_df_cols() function this time comparing the new combined dataset and 2017. 

```{r}
compare_df_cols(data_2015_2016, candy_2017, return = "mismatch", bind_method = "rbind")
```


7. From the comparison, I could see that all column titles included a question number at the beginning in the form q[1-9] which meant almost no column matched with my new dataframe. So I proceeded to clean the column names using clean_names(), then I employed the rename_with() function combined with gsub() allowed me to identify which column had the predetermined pattern and replace it with nothing hence "". I could then reuse compare_df_cols() which indicated a few columns were named similarly but needed correcting. 

```{r}
# pattern to identify question number at beginning of column name
pattern <- "q+[0-9]+_" 

candy_2017 <- candy_2017 %>% 
  # clean columns names
  clean_names() %>% 
  # rename all columns with this pattern, replace it with nothing
  rename_with(~ gsub(pattern, "", .x)) %>% 
  #rename common columns identified using comparison function
  rename(x100_grand_bar = "100_grand_bar", 
         state = "state_province_county_etc",
         brown_globs = "anonymous_brown_globs_that_come_in_black_and_orange_wrappers_a_k_a_mary_janes",
         bonkers = "bonkers_the_candy"
         )
```



8. I identifed some more columns in my combined column which when renamed would easily merge with the 2017 column upon binding them

```{r}
data_2015_2016 <- data_2015_2016 %>% 
  rename(
    country = "which_country_do_you_live_in",
    day = "which_day_do_you_prefer_friday_or_sunday",
    gender = "your_gender"
  )

```


9. I then combined both datasets into my final fully joined dataset

```{r}
combined_candy <- bind_rows(data_2015_2016, candy_2017)
```

10. Once created a cut all data that did not relate to candy using the irrelevant 
columns' index poisiton. I saved this new data to a new variable called, 
specific_candy_data. 

```{r}
specific_candy_data <- select(combined_candy, -18, -33, -38, -41, -45, -56, -63,
                              -88, -93:-95, -97:-112, -119, -130, -137:-139,
                              -145, -148:-156)
```


11. I rearranged the columns to bring the more important non-candy columns to
the front using select. I selected all the ones I wanted at the front, and then
used everything() to then include the rest. 

```{r}
specific_candy_data <- specific_candy_data %>% 
  # selected the columns i wanted at the front
  select(timestamp:going_out, day:which_state_province_county_do_you_live_in,
         internal_id:state, 
         #once columns selected everything() calls all the after columns afterwards
         everything())
```

12. I then changed the values of the timestamp and internal_id columns (which uniquely identified
each row) to represent the year. 

```{r}
specific_candy_data <- specific_candy_data %>% 
  # converted timestamp to character so perform regex function
  mutate(timestamp = as.character(timestamp)) %>% 
  # if string is "2015-+" then replace with 2015, else replace with "2016"
  mutate(timestamp = if_else(str_detect(timestamp, "2015-+"), "2015", "2016")) %>% 
  # repeat the same for the internal_id column - change to character
  mutate(internal_id = as.character(internal_id)) %>%  
  # if string is "90+" then replace with "2017", else replace with "NO YEAR"
  mutate(internal_id = if_else(str_detect(internal_id, "90+"), "2017", "NO YEAR"))
```

13. Now I can merge  both timestamp and internal_id together (using unite()) 
into a new column titled "year"

```{r}
specific_candy_data <-  specific_candy_data %>% 
  # merge together but to not include NAs
  unite("year", timestamp, internal_id, na.rm = TRUE) %>%
  # replace spaces left by not including NAs with NA
  mutate(year = na_if(year, ""))
```


14. The final column merge I did was to combine the state column that I had
missed when merging the two datasets. 

```{r}
specific_candy_data <-  specific_candy_data %>% 
  # merge together without NAs 
  unite("state_county", state, which_state_province_county_do_you_live_in, na.rm = TRUE) %>% 
  # replace spaces left by not including NAs with NA
  mutate(state_county = na_if(state_county, ""))
```

15. The age column proved difficult to clean. I used a combination of different
regex patterns which matched the variation of age entries in the dataset. 

```{r}
specific_candy_data <- specific_candy_data %>% 
  # the first pattern removes a specific data entry - 9.0E22
  mutate(age = str_replace_all(age, "\\.[0-9][a-zA-Z][0-9]", ""),
         # this removes any alphabetical entries
         age = str_replace_all(age, "[A-Za-z]+[A-Za-z]", ""),
         # this removes any punctuation followed by a number e.g. decimal points
         age = str_replace_all(age, "[:punct:][0-9]?", ""),
         # this pattern removes entries with an ellipsis
         age = str_replace_all(age, "...", ""),
         # replaces age == 0 with NA
         age = na_if(age, "0"))
```

16. The country column was the most difficult to clean effectively. I found it 
hard to create a singular expression to capture all the different spellings of 
USA/America etc. I managed with several regular expressions and some recoding of 
ones I could not get in my expressions. I saved this final piece of cleaning to 
a new dataframe titled 'final_candy'. 

```{r}
final_candy <- specific_candy_data %>%
   mutate(
     # removes punctuation
     country = str_replace_all(country, "[:punct:]", ""),
     # removes any numbers 
     country = str_replace_all(country, "[0-9]+", ""),
     # replaces different spellings of USA with USA
     country = str_replace_all(country, "[uU][sS][aA]? *[aA-zZ]?", "USA"),
     #replaces different spellings of United States with USA
     country = str_replace_all(country, "[uU][nN]i?t?ed?s? [sS]t?a?t?e?s?", "USA"),
     # replaces where country begins with a different spelling of USA with USA
     country = if_else(str_detect(country, "^[uU][sS][aA]"), "USA", country),
     # replaces with USA if different spellings of USA come at the end
     country = if_else(str_detect(country, "[uU][sS][aA]$"), "USA", country),
     # accounts for different spellings of Canada and replaces with "Canada"
     country = str_replace_all(country, "[cC][aA][nN][aA]?[dD]?[aA]?[eE]?+`?", "Canada"),
     # accounts for different spellings of UK - replaces with UK 
     country = str_replace_all(country, "[uU][nN][iI][tT][eE][dD] +[kK][iI][nN][gG]?[dD][oO][mM]", "UK"),
     # changes spellings of England to UK
     country = str_replace_all(country, "[eE]nd?g?land", "UK"),
     # accounts for case insensitivity for UK, replacing with "UK"
     country = str_replace_all(country, "[uU][kK]", "UK"),
     # I have re-coded ones that I could not get in my regular expressions 
     country = recode(country, 
                          "Murica" = "USA", 
                          "USA USA" = "USA",
                          "The Yoo Ess of Aaayyyyyy" = "USA", 
                          "USA of America" = "USA", 
                          "USAA" = "USA", 
                          "USASA USA" = "USA", 
                          "the best one USA" = "USA",
                          "USA think but its an election year so who can really tell" = "USA",
                          "America" = "USA", 
                          "Units States" = "USA", 
                          "USASA" = "USA", 
                          "SubCanadian North America Merica" = "USA",
                          "Merica" = "USA", 
                          "Trumpistan" = "USA", 
                          "USAtes" = "USA",
                          "america" = "USA", 
                          "USASA USASA" = "USA",
                          "United States of America" = "USA",
                          "The USA of America" = "USA", 
                          "unhinged states" = "USA",
                          "U S" = "USA", 
                          "North Carolina" = "USA", 
                          "Pittsburgh" = "USA",
                          "New York" = "USA", 
                          "California" = "USA", 
                          "New Jersey" = "USA", 
                          "Alaska" = "USA", 
                          "u s a" = "USA",
                          "AhemAmerca" = "USA",
                          "SubCanadaian North America Merica" = "USA",
                          "Not the USAr Canada" = "Canada",
                          "Scotland" = "UK",
                          "murrika" = "USA"),
     # if country is not either UK, Canada or USA, replace with "Other"
     country = if_else(country == c("USA", "UK", "Canada"), country, "Other")
         )
```

17. Finally, I wrote out my cleaned data as a csv to my clean data folder, ready
to be read in for analysis. 

```{r}
write_csv(final_candy, "clean_data/clean_candy_data.csv")
```


## Data Analysis

Read in my cleaned data

```{r}
clean_candy <- read_csv("clean_data/clean_candy_data.csv")
```

1. What is the total number of candy ratings given across the three years?

```{r}
clean_candy %>% 
   summarise(total_reviews = 
    sum(across( # too the sum across all columns of reviews 
      .cols = butterfinger:take_5, # first candy to last candy based on position
      .fns = ~sum(!is.na(.x)) # calculated the sum of each row not including NA
    )) # summed all sums together to get total
  )
```

2. What was the average age of people who are going out trick or treating?

```{r}
clean_candy %>% 
  select(age, going_out) %>% # select relevant columns
  filter(going_out == "Yes") %>% # filter where they are going out
  drop_na(age) %>% # drop NA values 
  summarise(avg_age = mean(age)) # return the average age
```


3. What was the average age of people who are not going trick or treating?

```{r}
clean_candy %>% 
  select(age, going_out) %>% # select columns
  filter(going_out == "No") %>% # filter where they are not going out
  drop_na(age) %>% # drop NA values
  summarise(avg_age = mean(age)) # return the average age
```


4. For each of joy, despair and meh, which candy bar received the most of these 
ratings?

```{r}
clean_candy %>% 
  #pivot all candy columns to "candy" and their values to "rating"
   pivot_longer(cols = butterfinger:take_5,
               names_to = "candy",
               values_to = "rating") %>% 
  select(candy, rating) %>% # select columns needed 
  group_by(candy, rating) %>% # group by both candy and rating 
  drop_na() %>% # drop NA values
  summarise(total = n()) %>% # num of ratings for candy bar based on rating
  ungroup() %>% # ungroup so to regroup 
  group_by(rating) %>% # group by the rating now 
  slice_max(total) # get the candy bar with the max total for each rating 
```


5. How many people rated Starburst as despair?

```{r}
clean_candy %>% 
  #pivot all candy columns to "candy" and their values to "rating"
   pivot_longer(cols = butterfinger:take_5,
               names_to = "candy",
               values_to = "rating") %>% 
  select(candy, rating) %>% # selecting columns needed
  filter(candy == "starburst") %>% # filter for starburst
  group_by(rating) %>% # group by rating 
  summarise(n = n()) %>% #total number of ratings per rating
  ungroup() %>% # ungroup 
  filter(rating == "DESPAIR") # return total for despair
```

6. What was the most popular candy bar by this rating system for each gender in 
the dataset ?
rating system = despair -1, joy +1, meh 0 

```{r}
clean_candy %>% 
  #pivot all candy columns to "candy" and their values to "rating"
   pivot_longer(cols = butterfinger:take_5,
               names_to = "candy",
               values_to = "rating") %>% 
  # case when used to create a new column which denotes score for rating
  mutate(rating_system = case_when(
    rating == "DESPAIR" ~ -1,
    rating == "JOY" ~ 1,
    rating == "MEH" ~ 0
  )) %>% 
  select(candy, gender, rating_system) %>% # select columns needed
  drop_na() %>% #drop NA values
  group_by(gender, candy) %>% # 
  summarise(total = sum(rating_system)) %>% # get total score for gender and candy
  ungroup() %>% #ungroup
  group_by(gender) %>% # group by gender to get total for each gender
  slice_max(total) # returns most popular based on rating for each gender
```


7. What was the most popular candy bar in each year?

```{r}
clean_candy %>% 
  # pivot all candy columns to "candy" and their values to "rating"
   pivot_longer(cols = butterfinger:take_5,
               names_to = "candy",
               values_to = "rating") %>% 
  # case when used to create a new column which denotes score for rating
  mutate(rating_system = case_when(
    rating == "DESPAIR" ~ -1,
    rating == "JOY" ~ 1,
    rating == "MEH" ~ 0
  )) %>% 
  select(year, candy, rating_system) %>% # select appropriate columns
  drop_na() %>% # drop NA values
  group_by(candy, year) %>% # group by candy and year
  summarise(total = sum(rating_system)) %>% # get the total for both based on rating
  ungroup() %>% # ungroup to carry out another group by 
  group_by(year) %>% # group by year 
  slice_max(total) # return most popular candy bar for each year
```

8. What was the most popular candy bar by this rating for people in US, Canada, 
 UK, and all other countries?

```{r}
clean_candy %>% 
  # pivot all candy columns to "candy" and their values to "rating"
   pivot_longer(cols = butterfinger:take_5,
               names_to = "candy",
               values_to = "rating") %>%
  # case when used to create a new column which denotes score for rating
  mutate(rating_system = case_when(
    rating == "DESPAIR" ~ -1,
    rating == "JOY" ~ 1,
    rating == "MEH" ~ 0
  )) %>% 
  select(country, candy, rating_system) %>% # select columns
  drop_na() %>% # drop NAs
  group_by(country, candy) %>% # group by country and candy 
  summarise(total = sum(rating_system)) %>% # sum of rating for country and candy
  ungroup() %>% # ungroup to group by again 
  group_by(country) %>% # group by country to get max total for each country
  slice_max(total) # return the most popular candy per country 
```



```{r}
```

